Parallel Hierarchical Clustering using \\ Rank-Two Nonnegative Matrix Factorization

### 2020-11-13 latex/paper.tex
briging down to 10 pages
added back in funding acks
adding description of rank scaling experiment, still need to add figure

\documentclass[conference,compsoc]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.

\usepackage{cite,doi}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithm,algpseudocode}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{cleveref}
\usepackage{tikz}
\usepackage{mathtools}
\usepackage{ourmacros}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

@@

\title{Parallel Hierarchical Clustering using \\ Rank-Two Nonnegative Matrix Factorization
\thanks{This material is based upon work supported by the National Science Foundation under Grant No. OAC-1642385 and OAC-1642410.
This manuscript has been co-authored by UT-Battelle, LLC under Contract No. DE-AC05-00OR22725 with the U.S. Department of
Energy. This project was partially funded by the Laboratory Director's Research and Development fund. This research used resources
of the Oak Ridge Leadership Computing Facility at the Oak Ridge National Laboratory, which is supported by the Office of Science of
the U.S. Department of Energy.}
}

@@

%\begin{IEEEkeywords}
%component, formatting, style, styling, insert
%\end{IEEEkeywords}

@@

\input{tex/ack.tex}

\bibliography{paper}
\bibliographystyle{plainurl}

>>>>

\documentclass[10pt, conference]{IEEEtran}
%\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.

\usepackage{cite,doi}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithm,algpseudocode}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{cleveref}
\usepackage{tikz}
\usepackage{mathtools}
\usepackage{ourmacros}
\usepackage{subcaption}

\usepackage[subtle,tracking=normal,leading=tight,leadingfraction=0.98, indent=tight, wordspacing=normal, bibliography=tight]{savetrees}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\usepackage{paralist}
\renewenvironment{itemize}[1]{\begin{compactitem}#1}{\end{compactitem}}

@@

\title{Parallel Hierarchical Clustering using \\ Rank-Two Nonnegative Matrix Factorization}

@@

\begin{IEEEkeywords}
low-rank approximation, distributed-memory parallel algorithms, scalable clustering
\end{IEEEkeywords}

@@

%Add this back in for camera ready
\input{tex/ack.tex}

\bibliographystyle{plainurl}
\bibliography{paper}

### 2020-11-13 latex/tex/experiments.tex
adding description of rank scaling experiment, still need to add figure

All the experiments in this section were conducted on Summit. Summit is a supercomputer created by IBM for the Oak Ridge National Laboratory. 
There are approximately 4,600 nodes on Summit. 
Each node contains two IBM POWER9 processors on separate sockets with 512 GB of DDR4 memory.
Each POWER9 processor utilizes 22 IBM SIMD Multi-Cores (SMCs), although one of these SMCs on each processor is dedicated to memory transfer and is therefore not available for computation. 
For node scaling experimeheight=2innts, all 42 available SMCs were utilized in each node so that every node computed with 42 separate MPI processes.
Additionally, every node also supports six NVIDIA Volta V100 accelerators but these were unused by our algorithm.

@@

>>>>

All the experiments in this section were conducted on Summit. Summit is a supercomputer created by IBM for the Oak Ridge National Laboratory. 
There are approximately 4,600 nodes on Summit. 
Each node contains two IBM POWER9 processors on separate sockets with 512 GB of DDR4 memory.
Each POWER9 processor utilizes 22 IBM SIMD Multi-Cores (SMCs), although one of these SMCs on each processor is dedicated to memory transfer and is therefore not available for computation. 
For node scaling experiments, all 42 available SMCs were utilized in each node so that every node computed with 42 separate MPI processes.
Additionally, every node also supports six NVIDIA Volta V100 accelerators but these were unused by our algorithm.

@@

\subsubsection{Rank Scaling}

To confirm the slow growth in running time of the hierarchical algorithm in terms of the number of clusters $k$, we perform rank scaling experiments for \hyper{} and synthetic data. 
Assuming a balanced tree and relatively small $k$, \Cref{eq:treecost} shows that the dominant computational cost is proportional to $\log k$, while flat NMF algorithms have a dominant cost that is linear in $k$ \cite{KBP17}.
\Cref{fig:rankscaling} shows the raw time for various values of $k$, confirming that running time for HierNMF grows more slowly in $k$ than a flat NMF algorithm from PLANC \cite{EH+19-TR} with the same number of columns and processor grid.

\begin{figure}
\caption{\GB{To be added...}}
\label{fig:rankscaling}
\end{figure}
